
```{r echo=F, warning=F, include=F}
# This is a code block
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
```
```{r echo = F, include = F}
system("ls ../input", intern=TRUE)
```

#### Hey there ! In this script we'd be creating some new features and then we'd take a look at how those relate with the outcome variable.

#### Along with that we'd take a look at a some variables which makes us raise an eyebrow because of some fishy values that they contain. We'd try to clean them too.


### Loading the code
```{r warning=F, echo=F, include=F}
library(readr)
library(dplyr)
library(caret)
library(Metrics)
library(ggplot2)

rus <- read_csv("../input/train.csv")
rus <- data.frame(rus)
```

### Dimensions of data
```{r warning=F, echo=F}
dim(rus)
```

### Structure of data
```{r warning=F, echo=F}

str(rus)

```

### Missing values
```{r warning=F, echo=F}
missing <- data.frame(sapply(rus, function(x) sum(is.na(x))*100/length(x)))
missing$feature <- names(rus)
missing$num <- c(1:nrow(missing))
colnames(missing) <- c("missing_ratio", "feature", "num")
missing <- missing[, c("num", "feature","missing_ratio")]
print(missing[missing$missing_ratio != 0,], row.names = F)

```

Total features with missing values
```{r warning=F, echo=F}
nrow(missing[missing$missing_ratio != 0,])

```


So lot of variables with missing values. 51 out of 292 features have missing data. Some of them have high percentage of missing values.

### Variables with almost zero variance

```{r warning = F, echo=F}
# zero variance variables

insignificant <- nearZeroVar(rus)
print(names(rus[ , insignificant]))

```

The above variables are not going to help in predictive models. It is better to remove them.

### Extracting date, month, year, weekday from timestamp

```{r warning = F, echo=F}
rus$date <- as.POSIXct(strptime(rus$timestamp, format = "%Y-%m-%d"))
rus$day <- as.integer(format(rus$date, "%d")) # day
rus$month <- as.factor(format(rus$date, "%m")) # month
rus$year <- as.integer(format(rus$date, "%Y")) # year
rus$weekday <- as.factor(format(rus$date, "%u")) # weekday
rus$yearmonth <- paste0(rus$year, rus$month)
rus$timestamp <- NULL
rus$date <- NULL


ggplot(data = rus, aes(x = as.factor(day), y = price_doc)) + geom_boxplot(fill = "#5C7457") + labs(title = "Date of the month vs Price", x = "Date", y = "Price")
ggplot(data = rus, aes(x = as.factor(month), y = price_doc)) + geom_boxplot(fill = "#EAC435") + labs(title = "Month vs Price", x = "Month", y = "Price")
ggplot(data = rus, aes(x = as.factor(year), y = price_doc)) + 
    geom_boxplot(fill = "#345995") +
    coord_cartesian(ylim = c(0,10000000)) + labs(title = "Year vs Price", x = "Year", y = "Price")

ggplot(data = rus, aes(x = as.factor(weekday), y = price_doc)) + geom_boxplot(fill = "#E40066") + labs(title = "Day of the week vs Price", x = "Day", y = "Price")

```

An upward trend with each passing year. But no trend with months, days or weekdays.

```{r warning = F, echo = F}
# max floor
rus$max_floor <- as.integer(rus$max_floor)

# number of rooms
rus$num_room <- as.integer(rus$num_room)

# kitchen area
rus$kitch_sq <- as.integer(rus$kitch_sq)

# floor
rus$floor_ratio <- rus$floor/rus$max_floor

```

## Creating new Features !

This involves cleaning some variables that are already present.

### Build year

Let us look at the distrbution table of Build_Year.
```{r warning=F, echo=F}
table(as.factor(rus$build_year))

```

Needs a little cleaning. I'm replacing the values values like 0,1,2 with NA


```{r warning = F}

rus$build_year <- as.character(rus$build_year)

rus$build_year[rus$build_year == "20052009"] <- "2005"
rus$build_year[rus$build_year == "0"] <- NA
rus$build_year[rus$build_year == "1"] <- NA
rus$build_year[rus$build_year == "20"] <- "2000"
rus$build_year[rus$build_year == "215"] <- "2015"
rus$build_year[rus$build_year == "3"] <- NA
rus$build_year[rus$build_year == "4965"] <- NA
rus$build_year[rus$build_year == "71"] <- NA

rus$build_year <- as.integer(rus$build_year)
```

### Let us create a new variable. Year_Old = 2020 - Build_Year

```{r warning=F, echo=F}
rus$year_old <- 2020 - rus$build_year
ggplot(data = subset(rus, rus$year_old < quantile(rus$year_old,0.99, na.rm = T) & rus$year_old > quantile(rus$year_old, 0.01, na.rm = T)), aes(x = year_old, y = price_doc)) + geom_jitter(alpha = 0.4, color = "#A2FAA3") + scale_x_continuous(breaks = seq(0,2000,50)) +
    labs(title = "House Age vs Price", x = "Years Old", y = "Price")

```


### Floor Inverse

The new variable is floor_inverse If a building has 100 floors and a specific house is on 40th floor then the floor_proportion would be 100 - 40 = 60

```{r warning=F, echo=F}
rus$floor_inverse <- as.integer(rus$max_floor) - as.integer(rus$floor)
ggplot(data = subset(rus, rus$floor_inverse > 0 & rus$floor_inverse < quantile(rus$floor_inverse, 0.99, na.rm = T)), aes( x = floor_inverse, y = price_doc)) + geom_jitter(alpha = 0.4, color = "#4F759B") +
    labs(title = "Floor(if building were upside down) vs Price", x = "Floor Inverse", y = "Price")


```

### Non-residential area to living area ratio

```{r warning=F, echo=F}
rus$non_living_area <- (rus$full_sq-rus$life_sq)*100/rus$full_sq

ggplot(data = subset(rus, rus$non_living_area > 0), aes(x = non_living_area, y = price_doc)) + geom_point(alpha = 0.4, color = "#1E555C") +
    labs(title = "Non Living Area vs Price", x = "Non Living Area", y = "Price")


```

### Area per Room

This is calculated by dividing total living area by number of rooms

```{r warning=F, echo=F}
rus$room_area <- rus$life_sq/rus$num_room
ggplot(data = subset(rus, rus$room_area < quantile(rus$room_area, 0.99, na.rm = T)), aes(x = room_area, y = price_doc)) + geom_jitter(alpha = 0.8, color = "#04724D") +
    labs(title = "Average Room Area", x = "Average Room Area", y = "Price")


```

### State of the apartment

```{r warning = F, echo=F}
table(as.factor(rus$state))
```

"33" seems to be an odd value.
So replace "33" with "3"

```{r warning=F, echo=F}
rus$state[rus$state == "33"] <- "3"
rus$state <- factor(rus$state)
ggplot(data = rus, aes(x = state, fill = state)) + geom_bar() +
    theme(legend.position = "none") +
    labs(title = "Apartment Condition vs Price", x = "State", y = "Price")


```

### ID features

Features which have ID in their name are not much of a use to predictive model. So it is better to remove them. All these features start with id_*

Following features have id in their names.
```{r warning = F, echo=F}
names(rus) <- tolower(names(rus))
id.vars <- grep("^[i][d]",names(rus))
print(names(rus[ , id.vars]))
# remove id variables
rus <- rus[ , -id.vars]

```

There is a lot of cleaning to be done. We can try to impute the missing values because some models can't handle missing values. I tried knn imputation but it did not improve the performance of XGBOOST model.

Stay tuned for more feature engineering.

### Please like the kernel. Your likes are my motivation. ;)


